List of Items

### 2021-03-16 manuscript.RmD
- update for NM&S submission

---
title             : "A Longitudinal Analysis of the Privacy Paradox" 
shorttitle        : "Longitudinal Analysis of the Privacy Paradox"
author:            
  - name          : "Tobias Dienlin"
    email         : "tobias.dienlin@uni-hohenheim.de"
    affiliation   : "1"
    address       : "University of Hohenheim, Department of Media Psychology (540F), 70599 Stuttgart, Germany"
    corresponding : yes
  - name          : "Philipp K. Masur"
    email         : "philipp.masur@uni-mainz.de"
    affiliation   : "2"
  - name          : "Sabine Trepte"
    email         : "sabine.trepte@uni-hohenheim.de"
    affiliation   : "1"
affiliation:

  - id            : "1"
    institution   : "University of Hohenheim"
  - id            : "2"
    institution   : "Johannes Gutenberg University Mainz"
author_note: |
  All authors contributed extensively to the work presented in this paper. TD, PM, & ST designed the study; PM supervised the data collection; PM administered the data importation; TD & PM wrote the code, ran the models, and analyzed the output data; TD wrote most parts of manuscript, and PM & ST contributed individual sections and comments; ST supervised the project and wrote the grant application (in 2012). 
  The authors declare no competing interests.
  This research was funded by the German Federal Ministry of Education and Research (BMBF) Grant 16KIS0094, awarded to Sabine Trepte.

  This manuscript features a companion website that includes detailed summaries of the statistical results, the code, additional analyses, and a reproducible version of the manuscript (https://tdienlin.github.io/privacy-paradox-longitudinal). The data can be downloaded from http://dx.doi.org/10.7802/1937. 
abstract: |
  The privacy paradox states that people's concerns about online privacy are unrelated to their online sharing of personal information. 
  On the basis of a representative sample of the German population, which includes `r knitr::load_cache(label = "load-data", object = "n_complete")` respondents who were interviewed at three waves separated by 6 months, we investigate the privacy paradox from a longitudinal perspective. Using a cross-lagged panel model with random intercepts, we differentiate between-person relations from within-person effects.
  Results revealed that people who were more concerned about their online privacy than others also shared slightly less personal information and had substantially more negative attitudes toward information sharing (between-person level). 
  People who were more concerned than usual also shared slightly less information than usual (within-person level). 
  We found no long-term effects of privacy concerns on information sharing or attitudes 6 months later.
  The results provide further evidence against the privacy paradox, but more research is needed to better understand the variables' potential causal relations.
keywords          : "privacy paradox, privacy concerns, information sharing, longitudinal analysis, structural equation modeling"
wordcount         : "`r wordcountaddin::word_count('manuscript.Rmd')`"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
bibliography      : 
  - [bibliography/references.bib]
  - [bibliography/r-references.bib]
lang              : "english"
class             : "man"
mask              : no
# to compile as pdf
# (a) move file "_site.yml" to folder "tmp"
# (b) uncomment the following lines until incl. "\raggedbottom"
# (c) comment lines after "output: html_document ..."
# output            : papaja::apa6_pdf
# header-includes:
#   - \setlength{\parskip}{0em}
#   - \raggedbottom
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
---

@@

```{r load-packs, include=F}
# Load and install packages
# devtools::install_github("tdienlin/td@v.0.0.2.5")  # uncommen to install
packages <- c("broman", "kableExtra", "knitr", "lavaan", "magrittr", "papaja"
             , "psych", "pwr", "td", "tidyverse", "semTools", "scales")
invisible(lapply(packages, library, character.only = T))

@@

# Introduction

@@

However, current research on the privacy paradox has one major limitation. 
To the best of our knowledge, most empirical studies conducted so far have investigated the privacy paradox from a between-person perspective.
By employing empirical tests of relations between people (e.g., cross-sectional questionnaires analyzed with multiple regression or Pearson correlations), studies have analyzed whether people who are more concerned than _others_ also share less personal information than _others_.
Although such a perspective is interesting and represents a viable first step, it cannot make informed claims regarding causality.
The privacy paradox, however, implies a causal perspective:
Does a person, if _he or she_ becomes more concerned about online privacy, then also share less personal information?
This mismatch is problematic because although between-person relations are, except for some special cases, a _necessary_ condition for causal within-person effects, they are by no means a _sufficient_ one.
For example, it could be that the between-person relation is determined other third variables. 
Hence, as the next step in investigating the privacy paradox and to better understand the causal relation between privacy concerns and information sharing, it is necessary to conduct studies with within-person designs.

With this study we aim to answer four major questions. 
First, on a between-person level, how are concerns about online privacy related to the online sharing of personal information? 
Second, on a within-person level, does information sharing decrease when concerns increase?
Third, what are the potential long-term effects?
Are changes in concerns related to changes in information sharing 6 months later and/or vice versa? 
Fourth, what is the role of privacy attitudes, do they mediate the relation between privacy concerns and information sharing?
To best answer and contextualize these questions, we first provide an in-depth theoretical analysis of the privacy paradox, after which we present the empirical results of a longitudinal panel study, which is representative of the German population.

@@

@baruhOnlinePrivacyConcerns2017 published the first empirical meta-analysis on the relations between privacy concerns and various forms of social media use (e.g., information sharing or SNS usage). 
On the basis of 37 studies, @baruhOnlinePrivacyConcerns2017 found a small and significant statistical relation between concerns about online privacy and online information sharing (_r_ = -.13, 95% CI [-.07, -.18]). 
Another more recent meta analysis by @yuMetaanalysisExplorePrivacy2020 also finds a significant bivariate relation between privacy concerns and information sharing, albeit smaller (_r_ = -.06, 95% CI [-.01, -.12]]). 
There also exist several systematic literature reviews on the privacy paradox [@barthPrivacyParadoxInvestigating2017; @gerberExplainingPrivacyParadox2018; @kokolakisPrivacyAttitudesPrivacy2017].
@kokolakisPrivacyAttitudesPrivacy2017 come to the conclusion that "the dichotomy between privacy attitude and behaviour should not be considered a paradox anymore." (p. 130)
However, the authors also note that the privacy paradox is a "complex phenomenon that has not been fully explained yet".
@barthPrivacyParadoxInvestigating2017 are more skeptical, and argue that "attempts to theoretically explain and practically solve the problem of the privacy paradox are still scarce and we feel the subject deserves far more research attention" (p. 1052).

<!-- In conclusion, large parts of the current literature suggests that a significant relation between concerns about online privacy and the online sharing of personal information exists and that it is small---which speaks against the privacy paradox. -->

@@

Privacy is defined as the "[...] voluntary and temporary withdrawal of a person from the general society through physical or psychological means [...]" [@westinPrivacyFreedom1967, 7]. 
Privacy captures aspects of both volitional _control_ and social _separateness_ [@braunlichLinkingLooseEnds2020; @marwickNetworkedPrivacyHow2014].
Several dimensions of privacy have been proposed.
For example, it is possible to distinguish a vertical and a horizontal level [@masurSituationalPrivacySelfdisclosure2018].
Whereas the vertical level captures privacy from authorities, institutions, or companies, horizontal privacy addresses privacy from peers, colleagues, or other people.
When it comes to concerns in general, interestingly they do not seem to be established as a stand-alone theoretical concept in psychology [@colmanDictionaryPsychology2015].
Lexically, concerns are defined as a "marked interest or regard usually arising through a personal tie or relationship" that also reflect an "uneasy state of blended interest, uncertainty, and apprehension" [@merriam-websterDefinitionConcern2018].
Concerns therefore represent both a latent _motivation_ (or increased attention), a negatively valenced _emotion_ (or affective condition), and are mostly implicit.
As a theoretical construct, privacy concerns can hence be categorized as an affective motivational disposition.
As such, there are many similarities with other concepts, including emotions (e.g., fear, anxiety), moods (e.g., dismay, fatigue), attitudes (risk perception, approval), values (e.g., autonomy, freedom), personality traits (e.g., introversion, risk avoidance), and even physiological activation (e.g., attention, arousal).
Taken together, concerns about online privacy represent how much an individual is motivated to focus on his or her control over a voluntary withdrawal from other people or societal institutions on the Internet, accompanied by an uneasy feeling that his or her privacy might be threatened.

The online sharing of personal information, on the other hand, captures how much person-related information people share when they use the Internet, including information about their age, sex, name, address, health, and finances.
Information sharing can be differentiated from communication and self-disclosure.
Communication is broad, because it comprises all verbal and nonverbal information that is emitted [e.g., @watzlawickPragmaticsHumanCommunication2011].
Self-disclosure is narrow, because it focuses on deliberate revelations about the true self to others [e.g., @jourardTransparentSelf1964].
Information sharing is even more specific, because it addresses only person-related information but ignores other types of self-disclosure such as personal fears, values, or plans.

@@

Currently, there is a lack of studies that explicitly analyze how behavior is affected by _concerns_ in general.
Fortunately, however, we know much about the behavioral effects of related concepts such as attitudes or fears, which all can affect behavior, sometimes profoundly [@fishbeinPredictingChangingBehavior2010; @rogersCognitivePhysiologicalProcesses1983].
Emotions, perhaps the concept most closely related to concerns, have a particularly strong effect on behavior. 
By causing fight, flight, or freeze reactions, they are a primordial trigger of behavior and are considered to be an adaptive mechanism of evolved species [@dolanEmotionCognitionBehavior2002].
<!-- With their direct link to the amygdala, emotions can trigger reactions already subcortically [i.e., without activation of the more recently evolved cortical structures; @dolanEmotionCognitionBehavior2002]. -->

Also empirically, concerns have been shown to affect behavior. 
People more concerned about the environment show more environment-related behaviors [@bambergHowDoesEnvironmental2003].
People more concerned about their appearance consume fewer calories [@hayesConcernAppearanceHealth1987].
People more concerned about their bodies engage in more physical exercise [@reelRelationsBodyConcerns2007].
Taken together, it is reasonable to expect that also concerns about online privacy should somehow reflect in the online sharing of personal information.

At the same time, there are some factors that likely diminish the relation.
Most prominently, there is the so-called _attitude behavior gap_ [@fishbeinPredictingChangingBehavior2010], which states that people sometimes act against their own attitudes. 
Evidently, not everyone concerned about their physical health exercises regularly.
The explanation is simple:
Other factors such as subjective norms and perceived behavioral control also determine behavior [@ajzenIntentionsActionsTheory1985], which automatically reduces the impact of attitudes or concerns. 
Specifically, two of the most influential factors that affect online information sharing are (a) strong subjective norms [@heirmanPredictingAdolescentsDisclosure2013] and (b) expected benefits [@krasnovaOnlineSocialNetworks2010]. 
In other words, users often prioritize social support, special offers, or improved services, accepting that their privacy will be diminished.
@trepteRiskyBehaviorsHow2014 listed several factors that can additionally attenuate the relation:
lack of strength of concerns,
absence of negative personal experiences, 
or situational constraints due to social desirability.
In conclusion, also in the context of the privacy paradox it _not_ reasonable to expect a _perfect_ relation between attitudes and behaviors.
However, we should still expect to find a relation that is _small_ or _moderate_.

@@

In conclusion, in line with prior research and the within-person rationales presented above, we expect to find a small significant relation between privacy concerns and information sharing, both on the between-person level and the within-person level.

@@

@@

In general, we believe that it should be possible to find long-term effects. 
It has been argued that privacy concerns affect privacy behavior in the long run [e.g., @heirmanPredictingAdolescentsDisclosure2013]. 
The underlying theoretical mechanism could be that the emotional part of privacy concerns causes (a) motivated information selection and (b) motivated information processing, which is likely to change actual behavior [@nabiCognitivefunctionalModelEffects1999]. 
Specifically, when privacy concerns increase (e.g., because of experienced or witnessed privacy infringements), people might begin reading more media articles on privacy issues and might also consume these articles more carefully, which could prompt information sharing practices that are more cautious.
Also empirically, a study with 290 participants found small negative longitudinal (between-person) relations between privacy concerns and self-disclosure [@koohikamaliInvestigationDynamicModel2019].

At the same time, the adverse effect seems plausible as well, with two potential outcomes. 
On the one hand, the long-term relation could be positive: 
If people start to share more information online, they might become increasingly aware that their privacy is at risk, thereby stirring concern [@tsay-vogelSocialMediaCultivating2018].
On the other hand, the long-term relation might also be negative:
When people share more personal information online they might become accustomed to doing so, which potentially reduces concern [for example, due to the mere exposure effect; @zajoncAttitudinalEffectsMere1968].
Finally, there could also be no long-term relation at all: 
People might have already become used to sharing information online, which stifles further cognitive or emotional processing.
This rationale is central to so-called _privacy cynicism_ [e.g., @hoffmannPrivacyCynicismNew2016].

@@

It has been argued that privacy attitudes could bridge the gap between concerns and information sharing [e.g., @dienlinPrivacyParadoxRelic2015]. 
In contrast to privacy concerns, privacy attitudes capture a more explicit, fluctuating cognitive appraisal [@tsay-vogelSocialMediaCultivating2018].
Although both variables are related to information disclosure, attitudes are likely the better predictor. 
This reasoning follows the rational choice paradigm [@simonBehavioralModelRational1955], which maintains that behavior is always at least partially influenced by convictions, attitudes, and cost-benefit analyses. 
Also empirically, a study of 1,042 youths from Belgium found that the relation between privacy attitudes and disclosure of personal information was strong (_r_ = .56), whereas the relation between privacy concerns and disclosure was only moderate [_r_ = -.29; @heirmanPredictingAdolescentsDisclosure2013].

@@

# Method
## Statistics

We follow the recommendation by @lakensJustifyYourAlpha2018 and first justify the choice of our alpha level. 
We determined adequate error margins by considering the potential implications of both false positive and false negative findings (i.e., alpha and beta errors): 
On the one hand, if we committed an alpha error, we would wrongfully conclude that people's concerns and behaviors are consistent. 
Communicating such a false result to the public might unjustly reassure people when they should be more alert.
On the other hand, if we committed a beta error, we would wrongfully conclude that individuals behave paradoxically.
Communicating such a false result would unjustly accuse people of implausible behavior, potentially causing unnecessary distress or reactance.
We consider both errors to be equally detrimental. 
Hence, we chose balanced error rates, setting a maximum error rate of 5% for both alpha and beta. 
As the smallest effect size of interest [SESOI; @lakensEquivalenceTestingPsychological2018], we chose to consider effects that are at least small [i.e., standardized coefficients above $\beta$ = `r sesoi %>% my_round("beta")`; @cohenPowerPrimer1992] as able to offer empirical support for our theoretical hypotheses.
Significantly smaller effects were not considered able to offer support.
The six hypotheses were tested with a one-tailed approach and the six research questions with a two-tailed approach.
On the basis of the balanced alpha-beta approach with a maximum error probability of `r alpha %>% round(3) * 100`%, a desired power of `r power %>% round(3) * 100`%, and an SESOI of $\beta$ = `r sesoi %>% my_round("beta")`, we calculated a minimum sample size of `r dollar_format(prefix = "")(round(n_apriori, 0))` respondents. 
Given the final sample size of `r dollar_format(prefix = "")(n)` respondents (see below), alpha and beta errors were balanced for our hypotheses (research questions) when we used a critical alpha of `r alpha_balanced_h  %>% round(3) * 100`% (`r alpha_balanced_rq  %>% round(3) * 100`%), resulting in a power of `r power_balanced_h %>% round(3) * 100`%  (`r power_balanced_rq %>% round(3) * 100`%) to detect small effects.

For the analyses, we used `r r_citations`.
The code, additional analyses, and a reproducible version of this manuscript can be found on the manuscript's companion website at https://tdienlin.github.io/privacy-paradox-longitudinal.

## Procedure and Respondents

This study is part of a large-scale project which investigates the development of privacy and self-disclosure, including several other variables. 
Other publications linked to the project can be accessed at https://osf.io/y35as/.
The data come from a longitudinal paper-and-pencil questionnaire study, in which a representative sample of the German population (16 years and older) was surveyed on overall five occasions.
The data can be downloaded from http://dx.doi.org/10.7802/1937.

@@

Privacy concerns were measured as a second-order factor. 
Three items captured the vertical dimension (e.g., "How concerned are you that institutions or intelligence services collect and analyze data that you disclosed on the Internet?"), and three items captured the horizontal dimension (e.g., "How concerned are you that people that you do not know might obtain information about you because of you online activities?"). 
Respondents rated all items on a 5-point scale ranging from 1 (*not at all concerned*) to 5 (*very concerned*). 
The means were _M_~t1~ = `r desc[["t1_pri_con_m_mean"]] %>% my_round(2)`, 
_M_~t2~ = `r desc[["t2_pri_con_m_mean"]] %>% my_round(2)`, 
_M_~t3~ = `r desc[["t3_pri_con_m_mean"]] %>% my_round(2)`, 
and the standard deviations _SD_~t1~ = `r desc[["t1_pri_con_m_sd"]] %>% my_round(2)`, 
_SD_~t2~ = `r desc[["t2_pri_con_m_sd"]] %>% my_round(2)`, and
_SD_~t3~ = `r desc[["t3_pri_con_m_sd"]] %>% my_round(2)`.
The two-dimensional model fit the data well, 
`r fit_txt(fit_pri_con_4)`.
The reliability was high ($\omega$~t1~ = `r reliabilityL2(fit_pri_con_4, "t1_pri_con_f")[["omegaL2"]] %>% my_round("std")`, $\omega$~t2~ = `r reliabilityL2(fit_pri_con_4, "t2_pri_con_f")[["omegaL2"]] %>% my_round("std")`, $\omega$~t3~ = `r reliabilityL2(fit_pri_con_4, "t3_pri_con_f")[["omegaL2"]] %>% my_round("std")`).
Overall, `r icc_pri_con$ICC_adjusted * 100 %>% round(0)`% of the measure's variance was explained by differences between persons.

@@

@@

The within-person results showed that when a person's privacy concerns increased, the same person also shared slightly less information online than usual.
Moreover, people who developed more positive attitudes toward the online sharing of personal information than usual, also shared substantially more personal information online.
Together, changes in concerns and attitudes are therefore related to changes in behavior, which speaks against the privacy paradox also on the within-person level.

@@

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

>>>>

---
title             : "A longitudinal analysis of the privacy paradox" 
shorttitle        : "A longitudinal analysis of the privacy paradox"
author:            
  - name          : "Tobias Dienlin"
    email         : "tobias.dienlin@univie.ac.at"
    affiliation   : "1"
    address       : "University of Vienna, Department of Communication, 1090 Vienna, Austria"
  - name          : "Philipp K. Masur"
    email         : "p.k.masur@vu.nl"
    affiliation   : "2"
    corresponding : yes
  - name          : "Sabine Trepte"
    email         : "sabine.trepte@uni-hohenheim.de"
    affiliation   : "3"
affiliation:
  - id            : "1"
    institution   : "University of Vienna"
  - id            : "2"
    institution   : "VU Amsterdam"
  - id            : "3"
    institution   : "University of Hohenheim"
authornote: |
  All authors contributed extensively to the work presented in this paper. TD, PM, & ST designed the study; PM supervised the data collection; PM administered the data importation; TD & PM wrote the code, ran the models, and analyzed the output data; TD wrote most parts of manuscript, and PM & ST contributed individual sections and comments; ST supervised the project and wrote the grant application (in 2012). 
  The authors declare no competing interests.
  This research was funded by the German Federal Ministry of Education and Research (BMBF) Grant 16KIS0094, awarded to Sabine Trepte.

  This manuscript features a companion website that includes detailed summaries of the statistical results, the code, additional analyses, and a reproducible version of the manuscript (https://tdienlin.github.io/privacy-paradox-longitudinal). The data can be downloaded from http://dx.doi.org/10.7802/1937

abstract: |
  The privacy paradox states that people’s concerns about online privacy are unrelated to their online sharing of personal information. On the basis of a representative sample of the German population, which includes `r knitr::load_cache(label = "load-data", object = "n_complete")` respondents interviewed at three waves separated by 6 months, we investigate the privacy paradox from a longitudinal perspective. Using a cross-lagged panel model with random intercepts, we differentiate between-person relations from within-person effects. Results revealed that people who were more concerned about their online privacy than others also shared slightly less personal information and had substantially more negative attitudes toward information sharing (between-person level). People who were more concerned than usual also shared slightly less information than usual (within-person level). We found no long-term effects of privacy concerns on information sharing or attitudes 6 months later. The results provide further evidence against the privacy paradox, but more research is needed to better understand potential causal relations.
keywords          : "privacy paradox, privacy concerns, information sharing, longitudinal analysis, structural equation modeling"
wordcount         : "`r wordcountaddin::word_count('manuscript.Rmd')`"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
class             : "man"
mask              : no
# to compile as pdf
# (a) move file "_site.yml" to folder "tmp"
# (b) uncomment the following lines until incl. "\raggedbottom"
# (c) comment lines after "output: html_document ..."
output            : papaja::apa6_pdf
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
# output:
#   html_document:
#     toc: TRUE
#     toc_float:
#       collapsed: FALSE
#     toc_depth: 2
bibliography      : 
  - [bibliography/references.bib]
  - [bibliography/r-references.bib]
csl               : "sage-harvard.csl"
---

@@

```{r load-packs, include=F}
# Load and install packages
# devtools::install_github("tdienlin/td@v.0.0.2.4")  # uncomment to install
# devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE) # uncomment to install
# devtools::install_github("crsh/papaja")  # uncomment to install
# devtools::install_github("cran/MissMech")
packages <- c("broman", "GGally", "kableExtra", "knitr", "lavaan", "magrittr", "papaja"
             , "psych", "pwr", "td", "tidyverse", "semTools", "scales")
invisible(lapply(packages, library, character.only = T))

@@

<!-- # Introduction -->

@@

However, current research on the privacy paradox has one major limitation. 
To the best of our knowledge, most empirical studies conducted so far have investigated the privacy paradox from a between-person perspective.
By employing empirical tests of relations between people (e.g., cross-sectional questionnaires analyzed with multiple regression or Pearson correlations), studies have analyzed whether people who are more concerned than _others_ also share less personal information than _others_.
Although such a perspective is interesting and represents a viable first step, it cannot make informed claims regarding causality.
The privacy paradox, however, implies a causal perspective:
Does a person, if _he or she_ becomes more concerned about online privacy, then also share less personal information?
This mismatch is problematic because although between-person relations are, except for some special cases, a _necessary_ condition for causal within-person effects, they are by no means a _sufficient_ one.
For example, it could be that the between-person relation is determined by other third variables. 
Hence, as the next step in investigating the privacy paradox and to better understand the causal relation between privacy concerns and information sharing, it is necessary to conduct studies with within-person designs.

With this study we aim to answer four major questions. 
First, on a between-person level, how are concerns about online privacy related to the online sharing of personal information? 
Second, on a within-person level, is information sharing lower than usual when concerns are higher than usual?
Third, what are the potential long-term effects?
Are changes in concerns related to changes in information sharing 6 months later and/or vice versa? 
Fourth, what is the role of privacy attitudes, do they mediate the relation between privacy concerns and information sharing?
To best answer and contextualize these questions, we first provide an in-depth theoretical analysis of the privacy paradox, after which we present the empirical results of a longitudinal panel study, which is representative of the German population.

@@

@baruhOnlinePrivacyConcerns2017 published the first empirical meta-analysis on the relations between privacy concerns and various forms of social media use (e.g., information sharing or SNS usage). 
On the basis of 37 studies, @baruhOnlinePrivacyConcerns2017 found a small and significant statistical relation between concerns about online privacy and online information sharing (_r_ = -.13, 95% CI [-.07, -.18]). 
Another more recent meta analysis by @yuMetaanalysisExplorePrivacy2020 also finds a significant bivariate relation between privacy concerns and information sharing, albeit smaller (_r_ = -.06, 95% CI [-.01, -.12]). 
There also exist several systematic literature reviews on the privacy paradox [@barthPrivacyParadoxInvestigating2017; @gerberExplainingPrivacyParadox2018; @kokolakisPrivacyAttitudesPrivacy2017].
@kokolakisPrivacyAttitudesPrivacy2017 come to the conclusion that "the dichotomy between privacy attitude and behaviour should not be considered a paradox anymore." (p. 130)
However, the authors also note that the privacy paradox is a "complex phenomenon that has not been fully explained yet".
@barthPrivacyParadoxInvestigating2017 are more skeptical, and argue that "attempts to theoretically explain and practically solve the problem of the privacy paradox are still scarce and we feel the subject deserves far more research attention" (p. 1052).

@@

Privacy is defined as the "[...] voluntary and temporary withdrawal of a person from the general society through physical or psychological means [...]" [@westinPrivacyFreedom1967, 7]. 
Privacy captures aspects of both volitional _control_ and social _separateness_ [@braunlichLinkingLooseEnds2020; @marwickNetworkedPrivacyHow2014].
People from all cultural backgrounds require privacy to fulfill fundamental needs including personal care, protected communication, intimacy, or sexuality [@altmanPrivacyRegulationCulturally1977; @westinPrivacyFreedom1967]. 
Being a universal human right [@assembly1948universal, Art. 12], privacy is essential for safety, psychosocial flourishing, and dignity.
It is driven by both individual needs and interpersonal negotiations thereof [@trepteSocialMediaPrivacy2020].

Several dimensions of privacy have been proposed.
For example, it is possible to distinguish a vertical and a horizontal level [@masurSituationalPrivacySelfdisclosure2018].
Whereas the vertical level captures privacy from authorities, institutions, or companies, horizontal privacy addresses privacy from peers, colleagues, or other people.
When it comes to concerns in general, interestingly they do not seem to be established as a stand-alone theoretical concept in psychology [@colmanDictionaryPsychology2015].
Concerns are usually understood as an uneasy mix of "interest, uncertainty, and apprehension" [@merriam-websterDefinitionConcern2018].
As a theoretical construct, privacy concerns can hence be categorized as an affective motivational disposition.
Taken together, concerns about online privacy represent how much an individual is motivated to focus on his or her control over a voluntary withdrawal from other people or societal institutions on the Internet, accompanied by an uneasy feeling that his or her privacy might be threatened.

The online sharing of personal information, on the other hand, captures how much person-related information people share when they use the Internet.
Information sharing can be differentiated from communication and self-disclosure.
Communication is broad, because it comprises all verbal and nonverbal information that is emitted [e.g., @watzlawickPragmaticsHumanCommunication2011].
Self-disclosure is more narrow, because it focuses on deliberate revelations about the true self to others, which including aspects such as personal fears, values, or plans [e.g., @jourardTransparentSelf1964].
Information sharing is even more specific, because it addresses only person-related information,  including information about their age, sex, name, address, health, and finances.

In what follows we hence investigate the two concepts of (a) concerns about online privacy and (b) online information sharing, aiming to investigate how they relate conceptually. In doing so, we adopt and focus on the perspective of individual people.

@@

Currently, there is a lack of studies that explicitly analyze how behavior is affected by _concerns_ in general.
Fortunately, however, we know much about the behavioral effects of related concepts such as attitudes or fears, which all can affect behavior, sometimes profoundly [@fishbeinPredictingChangingBehavior2010; @rogersCognitivePhysiologicalProcesses1983].
Emotions, perhaps the concept most closely related to concerns, have a particularly strong effect on behavior. 
By causing fight, flight, or freeze reactions, they are a primordial trigger of behavior and are considered to be an adaptive mechanism of evolved species [@dolanEmotionCognitionBehavior2002].

Also empirically, concerns have been shown to affect behavior [@hayesConcernAppearanceHealth1987; @reelRelationsBodyConcerns2007]. 
For example, people more concerned about the environment show more environment-related behaviors [@bambergHowDoesEnvironmental2003].
Taken together, it is reasonable to expect that also concerns about online privacy should somehow reflect in the online sharing of personal information.

At the same time, there are some factors that likely diminish the relation.
Most prominently, there is the so-called _attitude behavior gap_ [@fishbeinPredictingChangingBehavior2010], which states that people sometimes act against their own attitudes. 
Evidently, not everyone concerned about their physical health exercises regularly.
The explanation is simple:
Other factors such as subjective norms and perceived behavioral control also determine behavior [@ajzenIntentionsActionsTheory1985], which automatically reduces the impact of attitudes or concerns.

Specifically, two of the most influential factors that affect online information sharing are (a) strong subjective norms [@heirmanPredictingAdolescentsDisclosure2013] and (b) expected benefits [@krasnovaOnlineSocialNetworks2010]. 
In other words, users often prioritize social support, special offers, or improved services, accepting that their privacy will be diminished.
Sometimes, privacy concerns do not relate information sharing, because users lack the skills, knowledge, or literacy to change their online behavior, creating feelings of apathy or cynicism [@hargittaiWhatCanReally2016; @hoffmannPrivacyCynicismNew2016].
Likewise, personal information is also often shared by others, a phenomenon described as "networked privacy" [@marwickNetworkedPrivacyHow2014], which further reduces the power of individuals to determine how much personal information can be found online.
@trepteRiskyBehaviorsHow2014 listed several factors that can additionally attenuate the relation:
lack of strength of concerns,
absence of negative personal experiences, 
or situational constraints due to social desirability.
In conclusion, also in the context of the privacy paradox it is _not_ reasonable to expect a _perfect_ relation between attitudes and behaviors.
However, we should still expect to find a relation that is _small_ or _moderate_.

@@

In conclusion, we expect to find a small significant relation between privacy concerns and information sharing, both on the between-person level (Hypothesis 1) and the within-person level (Hypothesis 2).[^1]

@@

[^1]: To explain, with Hypothesis 1, we compare _different_ people with one another by analyzing their _average_ values across all measurements.
In other words, does a person, who is generally more concerned than others, also generally share less information than others?
With Hypothesis 2, we compare _specific_ measurements within the _same_ person.
In other words, does a person, if he/she is more concerned on T1 than on average, share more or less information on T1 than on average?

@@

In general, we believe that it should be possible to find long-term effects. 
It has been argued that privacy concerns affect privacy behavior in the long run [e.g., @heirmanPredictingAdolescentsDisclosure2013]. 
The underlying theoretical mechanism could be that the emotional part of privacy concerns causes (a) motivated information selection and (b) motivated information processing, which is likely to change actual behavior [@nabiCognitivefunctionalModelEffects1999]. 
Specifically, when privacy concerns are higher than usual (e.g., because of experienced or witnessed privacy infringements), people might begin reading more media articles on privacy issues and might also consume these articles more carefully, which could prompt information sharing practices that are more cautious.
Also empirically, a study with 290 participants found small negative longitudinal (between-person) relations between privacy concerns and self-disclosure [@koohikamaliInvestigationDynamicModel2019].

At the same time, the adverse effect seems plausible as well, with two potential outcomes. 
On the one hand, the long-term relation could be positive: 
If people start to share more information online, they might become increasingly aware that their privacy is at risk, thereby stirring concern [@tsay-vogelSocialMediaCultivating2018].
On the other hand, the long-term relation might also be negative:
When people share more personal information online they might become accustomed to doing so, which potentially reduces concern [for example, due to the mere exposure effect; @zajoncAttitudinalEffectsMere1968].
Finally, there could also be no long-term relation at all: 
People might have already become used to sharing information online, which stifles further cognitive or emotional processing.
This rationale is central to _privacy cynicism_ [e.g., @hoffmannPrivacyCynicismNew2016].

@@

It has been argued that privacy attitudes could bridge the gap between concerns and information sharing [e.g., @dienlinPrivacyParadoxRelic2015]. 
In contrast to general and implicit privacy concerns, privacy attitudes capture a more explicit, specific cognitive appraisal [@tsay-vogelSocialMediaCultivating2018]. 
Because general dispositions oftentimes affect more specific appraisals [@fishbeinPredictingChangingBehavior2010], general concerns about privacy may similarly affect more specific privacy attitudes [@dienlinPrivacyParadoxRelic2015].
This reasoning follows the rational choice paradigm [@simonBehavioralModelRational1955], which maintains that behavior is always at least partially influenced by specific convictions, attitudes, and cost-benefit analyses. 
Therefore, although both variables are related to information disclosure, attitudes are likely the better predictor. 
Also empirically, a study of 1,042 youths from Belgium found that the relation between privacy attitudes and disclosure intentions of personal information was strong (_r_ = .56), whereas the relation between privacy concerns and disclosure intentions was only moderate [_r_ = -.29; @heirmanPredictingAdolescentsDisclosure2013].

@@

# Method
## Procedure and Respondents

This study is part of a large-scale project which investigates the development of privacy and self-disclosure, including several other variables. 
Other publications linked to the project can be accessed at [link blinded during review].
The data come from a longitudinal paper-and-pencil questionnaire study, in which a representative sample of the German population (16 years and older) was surveyed on overall five occasions.
The data can be downloaded from [link blinded during review].

@@

Privacy concerns were measured as a second-order factor. 
Three self-developed items captured the vertical dimension (e.g., "How concerned are you that institutions or intelligence services collect and analyze data that you disclosed on the Internet?"), and three items by @buchananDevelopmentMeasuresOnline2007 captured the horizontal dimension (e.g., "How concerned are you that people that you do not know might obtain information about you because of you online activities?"). 
Respondents rated all items on a 5-point scale ranging from 1 (*not at all concerned*) to 5 (*very concerned*). 
The means were _M_~t1~ = `r desc[["t1_pri_con_m_mean"]] %>% my_round(2)`, 
_M_~t2~ = `r desc[["t2_pri_con_m_mean"]] %>% my_round(2)`, 
_M_~t3~ = `r desc[["t3_pri_con_m_mean"]] %>% my_round(2)`, 
and the standard deviations _SD_~t1~ = `r desc[["t1_pri_con_m_sd"]] %>% my_round(2)`, 
_SD_~t2~ = `r desc[["t2_pri_con_m_sd"]] %>% my_round(2)`, and
_SD_~t3~ = `r desc[["t3_pri_con_m_sd"]] %>% my_round(2)`.
The two-dimensional model fit the data well, 
`r fit_txt(fit_pri_con_4)`.
The reliability was high ($\omega$~t1~ = `r reliabilityL2(fit_pri_con_4, "t1_pri_con_f")[["omegaL2"]] %>% my_round("std")`, $\omega$~t2~ = `r reliabilityL2(fit_pri_con_4, "t2_pri_con_f")[["omegaL2"]] %>% my_round("std")`, $\omega$~t3~ = `r reliabilityL2(fit_pri_con_4, "t3_pri_con_f")[["omegaL2"]] %>% my_round("std")`).
Overall, `r icc_pri_con$ICC_adjusted * 100 %>% round(0)`% of the measure's variance was explained by differences between persons.

@@

In an additional analysis, we also tested the same model with a 1 year interval, which allowed to include data spanning until winter 2016 and 2017.
Most effects remained the same.
For example, we again found that people more concerned than others were less positive regarding information sharing (_r_ = -.36, _p_ < .001 ) and shared less information (_r_ = -.15, _p_ = .002). 
Likewise, people more positive toward data sharing than others also shared substantially more data (_r_ = .66, _p_ < .001).
Because including these two additional waves significantly reduces sample size, and because we consider it more likely that effects take place more immediately, these results should be considered exploratory.
For an overview of the results, see the additional analyses on our companion website (Section 2.1.2.7).

@@

The within-person results showed that when a person's privacy concerns are higher than usual, the same person also shared slightly less information online than usual.
Moreover, people who developed more positive attitudes toward the online sharing of personal information than usual, also shared substantially more personal information online.
Together, changes in concerns and attitudes are therefore related to changes in behavior, which speaks against the privacy paradox also on the within-person level.

@@
